---
title: "Livestock abortions reporting"
author: "Abel"
date: "03 october 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
# 04th October 2023
#Reading the file document
library(readxl)
Abortions_all_03102023 <- read_excel("~/Documents/R/Reporting of livestock abortions/Abortions_all_03102023.xlsx")
View(Abortions_all_03102023)
```
```{r}
## Loading packages
pacman::p_load(sf,tidyterra,dplyr)
library(tidyverse)
library(arsenal)
library(ggplot2)
library(lubridate)
library(broom)
library(data.table)
library(here)
library(rio)
library(tidyverse)
library(linelist)
library(janitor)
library(knitr)
library(finalfit)
library(forestmodel)
library(kableExtra)
library(skimr)
library(DescTools)
library(lmtest)
library("epitools")

```
```{r}
## Univariate analysis
table(Abortions_all_03102023$rvfelisa_IgM)
table(Abortions_all_03102023$rvfelisa_IgG)
table(Abortions_all_03102023$host_rvf)
table(Abortions_all_03102023$host_brucellosis)
table(Abortions_all_03102023$gender)
table(Abortions_all_03102023$occupation)
table(Abortions_all_03102023$education)
table(Abortions_all_03102023$subcounty)
table(Abortions_all_03102023$animalclinical_presentations)
table(Abortions_all_03102023$stage_pregnancy)
table(Abortions_all_03102023$environmental_features)
table(Abortions_all_03102023$history_animal_movement)
table(Abortions_all_03102023$rvf_vaccination_status)
# Calculating the median and IQR for the household
# Remove rows with missing values
cleaned_data <- na.omit(Abortions_all_03102023$household_size)

# Calculate the median of household_size from the cleaned vector
median_value <- median(cleaned_data)

# Calculate the IQR of household_size from the cleaned vector
iqr_value <- IQR(cleaned_data)

# Print the results
cat("Median:", median_value, "\n")
cat("IQR:", iqr_value, "\n")
```

```{r}
## Bivariate analysis
install.packages("arsenal")
library(arsenal)
Abortion2 <- tableby(rvfelisa_IgM ~ host_rvf + subcounty + rvf_vaccination_status + stage_pregnancy + environmental_features + history_animal_movement, data = Abortions_all_03102023)
summary(Abortion2, text = TRUE)

## Ploting for the results: 25th October 2023
IgG_bivariate <- tableby(rvfelisa_IgG ~ host_rvf, data = Abortions_all_03102023)
summary(IgG_bivariate, text = TRUE)

IgM_bivariate <- tableby(rvfelisa_IgM ~ host_rvf, data = Abortions_all_03102023)
summary(IgM_bivariate, text = TRUE)

#Ploting for brucellosis
brucella<- tableby(brucellosis ~ host_brucellosis, data = Abortions_all_03102023)
summary(brucella, text = TRUE)

#Understanding hosts that suffered both brucellosis and RVF

# Subset the data for hosts with both conditions
hosts_with_both <- Abortions_all_03102023 %>%
  filter(brucellosis == "Positive" & rvfelisa_IgG == "Positive")

# Count how many of each host type have both conditions
count_hosts <- table(hosts_with_both$host_rvf)

# Print the results
print(count_hosts)

# Create a combined variable that indicates the presence of both diseases, one of them, or neither
Abortions_all_03102023$both_diseases <- with(Abortions_all_03102023,
                                             ifelse(brucellosis == "Positive" & rvfelisa_IgG == "Positive", "Both",
                                                    ifelse(brucellosis == "Positive", "Only Brucellosis",
                                                           ifelse(rvfelisa_IgG == "Positive", "Only RVF IgG", "Neither"))))

# Create a contingency table
disease_table <- table(Abortions_all_03102023$host_rvf, Abortions_all_03102023$both_diseases)

# Print the results
print(disease_table)

```
```{r}
#9th October 2023, 2nd February 2024
# Multivariate analysis
library(dplyr)
log_analysis <- Abortions_all_03102023 %>%
  select(date_alert,rvfelisa_IgG) %>% 
  na.omit() %>% 
  mutate(rvfelisa_IgG = if_else(rvfelisa_IgG == "Positive", 1, 0))
logistic.display <- function(model) {
cat("Coefficients:\n")
print(m$coefficients[, 1:4, drop = FALSE])
cat("\n")
cat("Goodness of Fit:\n")
print(c("Deviance = ", round(m$deviance, 3)))
print(c("AIC = ", round(m$aic, 3)))
print(c("BIC = ", round(m$bic, 3)))
}
#count unique values for each variable
sapply(lapply(Abortions_all_03102023, unique), length)
#fiting the positives to be 1 and negatives to be 0
Abortions_all_03102023$rvfelisa_IgG_bin <- ifelse(Abortions_all_03102023$rvfelisa_IgG == "Positive", 1, 0)
# fiting the logistic regression model
Abortion4 <- glm(rvfelisa_IgG_bin ~ host_rvf + stage_pregnancy + environmental_features + history_animal_movement, family = binomial, data = Abortions_all_03102023)
summary(Abortion4)
,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,
##Simplifying the tables in R and making the model outputs into a publishable version
install.packages("gtsummary")
library(gtsummary)
Abortion4  %>% 
tbl_regression(exponentiate = TRUE) %>% 
bold_p()
,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,,
```{r}
```
```{r}
#date and IgM 
#command shift M piping
#Grouping number of abortions per day
IgGresult <- abortions5 %>% select(date_alert,rvfelisa_IgG) %>% na.omit() %>% 
  mutate(rvfelisa_num = if_else(rvfelisa_IgG == "Positive", 1, 0)) %>% 
  group_by(date_alert) %>% 
  summarise(number_reported = n())
IgGresult

#labelling the epicurve
library(dplyr)
library(ggplot2)
# Check if the dates in events are in IgGresult
print(filter(IgGresult, !is.na(event)))
ggplot(IgGresult, aes(x = date_alert, y = number_reported)) +
  geom_line() +
  geom_point() + 
  geom_text(data = subset(IgGresult, !is.na(event)),
            aes(label = event), nudge_y = 5) +  # Nudging the label upwards to avoid overlap
  labs(title = "Number of RVF ELISA IgG Positive Cases Over Time", 
       x = "Date", 
       y = "Number Reported") +
  theme_minimal()
+ ylim(0, max(IgGresult$number_reported) + 10)  # Adjusting the y-axis limits
  
```
```{r}
ggplot(IgGresult,aes(x = date_alert, y = number_reported)) +
  geom_line() + geom_point() + theme_classic()
```
```{r}
# Ploting the spatial distributions of the reported cases
library(pacman)
pacman::p_load(sf,ggspatial)
IgGreporting_locations <- abortions5 %>% 
  select(date_alert,rvfelisa_IgG,`gps_coordinates _latitude`,`gps_coordinates _longitude`) %>% 
  rename(latitude = `gps_coordinates _latitude`,longitude = `gps_coordinates _longitude`) %>% na.omit() %>% 
  st_as_sf(coords = c("longitude", "latitude"), crs = 4236)

#Isingiro_shape <- geodata::gadm("uganda", level = 3, path = "uganda")
```
```{r}
ggplot(IgGreporting_locations) + 
  geom_sf(aes(colour = rvfelisa_IgG))
```
```{r}
# Load necessary libraries
library(ggplot2)
library(terra)
library(sf)
library(ggspatial)

# Load and preprocess Isingiro shape
isingiro_shape <- terra::vect("uga_admbnda_ubos_20200824_shp/uga_admbnda_adm3_ubos_20200824.shp")
isingiro_shape <- isingiro_shape[isingiro_shape$ADM3_EN == "Isingiro"]

# Convert SpatVector to sf
isingiro_sf <- sf::st_as_sf(isingiro_shape)

# Plot the Isingiro shape (assuming it's now an sf object)
ggplot() +
  geom_sf(data = isingiro_sf, fill = NA)

# If you have another dataset named IgGreporting_locations that you'd like to overlay on the Isingiro shape:
ggplot() + 
  geom_sf(data = isingiro_sf, fill = NA) +
  geom_sf(data = IgGreporting_locations, aes(colour = rvfelisa_IgG))

# Ploting for different regions in Uganda
region_shape <- terra::vect("uga_admbnda_ubos_20200824_shp/uga_admbnda_adm2_ubos_20200824.shp")
region_shape <- uganda_shape[uganda_shape$ADM2_EN == "Regions"]

# Convert SpatVector to sf
region_shape<- sf::st_as_sf(region_shape)

# Plot the Isingiro shape (assuming it's now an sf object)
ggplot() +
  geom_sf(data = region_shape, fill = NA)

# If you have another dataset named IgGreporting_locations that you'd like to overlay on the Isingiro shape:
ggplot() + 
  geom_sf(data = region_shape, fill = NA) +
  geom_sf(data = IgGreporting_locations, aes(colour = rvfelisa_IgG))



## ploting a map of Uganda showing our distribution

# Load and preprocess Isingiro shape
uganda_shape <- terra::vect("uga_admbnda_ubos_20200824_shp")

# Convert SpatVector to sf
uganda_shape <- sf::st_as_sf(uganda_shape)

# Plot the Isingiro shape (assuming it's now an sf object)
ggplot() +
  geom_sf(data = uganda_shape, fill = NA)

# If you have another dataset named IgGreporting_locations that you'd like to overlay on the Isingiro shape:
ggplot() + 
  geom_sf(data = uganda_shape, fill = NA) +
  geom_sf(data = IgGreporting_locations, aes(colour = rvfelisa_IgG)
```
```{r}
#17th October 2023
#Understanding the duration difference between date_alert and date_sample collection
Abortions_all_03102023[["date_alert"]] <- as.Date(Abortions_all_03102023[["date_alert"]], format = "%Y-%m-%d") 
Abortions_all_03102023[["date _samplecollection"]] <- as.Date(Abortions_all_03102023[["date _samplecollection"]], format = "%Y-%m-%d")
Abortions_all_03102023$days_difference <- as.numeric(Abortions_all_03102023[["date _samplecollection"]] - Abortions_all_03102023[["date_alert"]])
average_difference <- mean(Abortions_all_03102023$days_difference, na.rm = TRUE) # Getting the difference between the dates
print(average_difference)

# Plotting the differences
library(ggplot2)
ggplot(Abortions_all_03102023, aes(x = days_difference)) +
  geom_histogram(binwidth=1, fill="steelblue", color="black", alpha=0.7) +
  labs(title = "Distribution of Days Difference between Alert and Sample Collection",
       x = "Days Difference",
       y = "Count") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))


```
```{r}
# Further analysis with Dr Andrew Conlan (17th October 2023)
#Loading packages
require(tidyverse)
library(readxl)

df <- read_excel("~/Documents/R/Reporting of livestock abortions/Abortions_all_03102023.xlsx")

df <- df %>% drop_na(rvfelisa_IgG)
df <- df %>% mutate(rvfelisa_IgG = rvfelisa_IgG=='Positive')
glmfit <- glm(rvfelisa_IgG  ~ host_rvf + 
                stage_pregnancy + 
                environmental_features + 
                history_animal_movement, data = df,family='binomial')
summary(glmfit)
## Running the time difference between occurence of abortion and sample collection
df %>% select(date_alert,`date _samplecollection`) %>% 
  mutate(difference = days(`date _samplecollection` - date_alert),.before=everything())
## Estimating the odds ratio from the logistic regression
glmfit2 <- glm(rvfelisa_IgG  ~ host_rvf + 
                environmental_features + 
                history_animal_movement, data = df,family='binomial')
summary(glmfit2)

# Round the odds ratios and their confidence intervals to 3 decimal places
# Compute the odds ratios
odds_ratios <- exp(coef(glmfit2))

# Compute the 95% confidence intervals for the odds ratios
conf_int <- confint(glmfit2)
odds_ratios_ci <- exp(conf_int)

# Round the odds ratios and their confidence intervals to 3 decimal places
rounded_odds_ratios <- round(odds_ratios, 3)
rounded_odds_ratios_ci <- round(odds_ratios_ci, 3)

# Display the rounded odds ratios and their confidence intervals
data.frame(OddsRatio = rounded_odds_ratios,
           CI_Lower = rounded_odds_ratios_ci[, 1],
           CI_Upper = rounded_odds_ratios_ci[, 2])

glmfit2 <- glm(rvfelisa_IgG ~ host_rvf + environmental_features + history_animal_movement, data = df, family = 'binomial')
# Compute the odds ratios
odds_ratios2 <- exp(coef(glmfit2))

# Compute the 95% confidence intervals for the odds ratios
conf_int2 <- confint(glmfit2)
odds_ratios_ci2 <- exp(conf_int2)

# Round the odds ratios and their confidence intervals to 3 decimal places
rounded_odds_ratios2 <- round(odds_ratios2, 3)
rounded_odds_ratios_ci2 <- round(odds_ratios_ci2, 3)

# Create and print the rounded odds ratios and their confidence intervals
results <- data.frame(OddsRatio = rounded_odds_ratios2,
                      CI_Lower = rounded_odds_ratios_ci2[, 1],
                      CI_Upper = rounded_odds_ratios_ci2[, 2])

print(results)

# Compute the odds ratios
odds_ratios2 <- exp(coef(glmfit2))

# Compute the 95% confidence intervals for the odds ratios
conf_int2 <- confint(glmfit2)
odds_ratios_ci2 <- exp(conf_int2)

# Extract p-values for the coefficients
glm_summary <- summary(glmfit2)
p_values <- glm_summary$coefficients[, 4]

# Round the odds ratios, their confidence intervals, and p-values to 3 decimal places
rounded_odds_ratios2 <- round(odds_ratios2, 3)
rounded_odds_ratios_ci2 <- round(odds_ratios_ci2, 3)
rounded_p_values <- round(p_values, 3)

# Create and print the table
results <- data.frame(OddsRatio = rounded_odds_ratios2,
                      CI_Lower = rounded_odds_ratios_ci2[, 1],
                      CI_Upper = rounded_odds_ratios_ci2[, 2],
                      P_Value = rounded_p_values)

print(results)


df <- df %>% select(date_alert,`date _samplecollection`) %>% 
  mutate(difference = as.numeric(`date _samplecollection` - date_alert)/(24*60*60),.before=everything())
```
```{r}
#Further analysis for call centre data (25th October 2023)
library(readxl)
data_callcentre <- read_excel("data_callcentre.xlsx")
View(data_callcentre)
```
```{r}
#Univariate analysis of the key variables
table(data_callcentre$reporting_platform)
table(data_callcentre$breed)
table(data_callcentre$number_abortions)
names(data_callcentre)
## number of abortions reported per herd
abortionsreported <- na.omit(data_callcentre$number_abortions)

# Calculate the median of household_size from the cleaned vector
median_value1 <- median(abortionsreported)

# Calculate the IQR of household_size from the cleaned vector
iqr_value <- IQR(abortionsreported )

# Print the results
cat("Median:", median_value1, "\n")
cat("IQR:", iqr_value, "\n")
## estimating the average parity of the animals on the farm

library(readxl)
data_callcentre <- read_excel("data_callcentre.xlsx")
View(data_callcentre)
head(data_callcentre$`age_host (weeks)`)
# See the unique values and how many times each appears
print(table(data_callcentre$`age_host (weeks)`))

# Check the structure of the variable
str(data_callcentre$`age_host (weeks)`)

# Check the first few rows
head(data_callcentre)
options(max.print = 1000)  # Adjust the number as necessary
print(data_callcentre$`age_host (weeks)`)
# Calculate mean and median
sum(is.na(data_callcentre$`age_host (weeks)`))
data_callcentre$`age_host (weeks)` <- as.numeric(as.character(data_callcentre$`age_host (weeks)`))
mean_value <- mean(data_callcentre$`age_host (weeks)`, na.rm = TRUE)
median_value <- median(data_callcentre$`age_host (weeks)`, na.rm = TRUE)

# Print the results
cat("Mean:", mean_value, "\n")
cat("Median:", median_value, "\n")
iqr_value <- IQR(median_value)
# Computing IQR for the parity
# Calculate the IQR of age_host (weeks) from your dataset
parityhost <- data_callcentre$`age_host (weeks)`

# Calculate the IQR
iqr_parityhost <- IQR(parityhost, na.rm = TRUE)

# Print the result
cat("IQR:", iqr_parityhost, "\n")

## Standard deviation

# Extract the 'age_host (weeks)' column from your dataset
parityhost <- data_callcentre$`age_host (weeks)`

# Calculate the mean
mean_parityhost <- mean(parityhost, na.rm = TRUE)

# Calculate the standard deviation
sd_parityhost <- sd(parityhost, na.rm = TRUE)

# Print the results
cat("Mean:", mean_parityhost, "\n")
cat("Standard Deviation:", sd_parityhost, "\n")

#Ploting the prevalence of RVF over time, 06th November 2023

library(readxl)
library(dplyr)
library(ggplot2)
library(tidyverse)
# Load the data
df <- read_excel("~/Documents/R/Reporting of livestock abortions/Abortions_all_03102023.xlsx")

# Convert rvfelisa_IgG to a logical vector (TRUE if Positive, FALSE if not)
df <- df %>% mutate(rvfelisa_IgG = rvfelisa_IgG == 'Positive')

# Ensure that the date_alert column is a Date type
df$date_alert <- as.Date(df$date_alert)

# Now, plot the data

## Ploting for both positives and negatives from IgG on the same map
# First, make sure 'rvfelisa_IgG' is a factor with the levels 'Positive' and 'Negative'
df$rvfelisa_IgG <- factor(df$rvfelisa_IgG, levels = c(FALSE, TRUE), labels = c("Negative", "Positive"))

# Now, aggregate to get the count of positives and negatives per date
df_summary <- df %>%
  group_by(date_alert, rvfelisa_IgG) %>%
  summarize(count = n(), .groups = 'drop') # count the number of occurrences

# Plot the data
ggplot(df_summary, aes(x = date_alert, y = count, fill = rvfelisa_IgG)) +
  geom_col() +  # Use geom_col() to create stacked bars
  scale_fill_manual(values = c("Negative" = "green", "Positive" = "red")) +  # Color code
  labs(x = "Date of Alert", y = "Count of IgG Results", title = "Proportion of IgG Results Over Time") +
  theme_minimal() +
  theme(legend.title = element_blank()) # Remove the legend title if desired

library(ggplot2)
library(dplyr)
library(scales)  # For date formatting

# Assuming your dataframe is df and it has been read and pre-processed

# Ensure that date_alert is a Date object
df$date_alert <- as.Date(df$date_alert, format = "%B %Y") # Adjust format based on your date structure

# Make sure that rvfelisa_IgG is a factor and has the levels set correctly
df$rvfelisa_IgG <- factor(df$rvfelisa_IgG, levels = c("Negative", "Positive"))

# Aggregate to get the count of positives and negatives per date
df_summary <- df %>%
  group_by(date_alert, rvfelisa_IgG) %>%
  summarize(count = n(), .groups = 'drop') # Count the number of occurrences

# Plotting the data
ggplot(df_summary, aes(x = date_alert, y = count, fill = rvfelisa_IgG)) +
  geom_col(position = "dodge") + # Use geom_col() to create bars side by side (dodged)
  scale_fill_manual(values = c("Negative" = "green", "Positive" = "red")) + # Assign colors
  scale_x_date(date_labels = "%b '%y", date_breaks = "1 month") + # Custom date format for x-axis
  labs(x = "Date of Alert", y = "Count of IgG Results", title = "Proportion of IgG Results Over Time") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotate x labels for better readability

# Ploting a bar graph if IgG and Date_alert 
library(readxl)
library(dplyr)
library(ggplot2)
library(scales)
library(lubridate)

# Loading the data
df <- read_excel("~/Documents/R/Reporting of livestock abortions/Abortions_all_03102023.xlsx")

# Converting 'date_alert' to Date type (if it's not already)
df$date_alert <- as.Date(df$date_alert)

# Extracting year and month for grouping
df <- df %>%
  mutate(year_month = floor_date(date_alert, "month"))

# Counting the positives and negatives per month
monthly_counts <- df %>%
  group_by(year_month, rvfelisa_IgG) %>%
  summarise(count = n(), .groups = 'drop')

# Plotting the bar graph
ggplot(monthly_counts, aes(x = year_month, y = count, fill = rvfelisa_IgG)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("Positive" = "red", "Negative" = "green")) +
  scale_x_date(date_breaks = "1 month", labels = date_format("%b %Y")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Date Alert", y = "Count of IgG Results", title = "Monthly IgG Results Over Time")

#Ploting the Epicurve/Histogram for IgG and Date_Alert
library(readxl)
library(dplyr)
library(ggplot2)
library(scales)
library(lubridate)

# Loading the data
df <- read_excel("~/Documents/R/Reporting of livestock abortions/Abortions_all_03102023.xlsx")

# Convert 'date_alert' to Date type (if it's not already)
df$date_alert <- as.Date(df$date_alert)

# Creating a new column that marks the positive results
df <- df %>% 
  mutate(positive = ifelse(rvfelisa_IgG == "Positive", 1, 0),
         month_year = floor_date(date_alert, "month"))

# Plotting the Time series for Reported Abortions
library(ggplot2)
library(scales)  # Make sure to load the 'scales' library for date formatting

ggplot(df, aes(x = month_year, fill = rvfelisa_IgG)) +
  geom_histogram(stat = "count", position = "stack", binwidth = 30) +
  scale_fill_manual(values = c("Positive" = "red", "Negative" = "green")) +
  scale_x_date(labels = date_format("%b %Y"), date_breaks = "1 month") +  # Corrected date formatting
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Date (Month/Year)", y = "Number of reported abortions", title = "Time series for Reported Abortions") +
  guides(fill = guide_legend(title = "RVF ELISA IgG"))

```
```{r}
## Shape files for Isingiro District

# Install necessary packages if you haven't already
install.packages("sf")
install.packages("ggplot2")
install.packages("dplyr")

# Load the packages
library(sf)
library(ggplot2)
library(dplyr)

# Read the shapefile
shapefile_path <- "~/Documents/R/Reporting of livestock abortions/abortions reporting/uga_admbnda_ubos_20200824_shp/uga_admbnda_adm3_ubos_20200824.shp"
uganda_map <- st_read(shapefile_path)

# If the shapefile contains many districts and you only want Isingiro, filter for it
isingiro_district <- uganda_map %>% 
  filter(ADM3_EN == "Isingiro") # Make sure 'ADM3_EN' is the correct field name for district names

# Read the dataset with IgG results and GPS coordinates
data_path <- "~/Abel Wilson Walekhwa Dropbox/Abel Wilson Walekhwa/Mac/Documents/R/Reporting of livestock abortions/Abortions reporting/mpox_27112023.xlsx"
library(readxl)
mpox_data <- read_excel(data_path)

# Replacing these with the actual column names from your dataframe
longitude_column_name <- "gps_coordinates _longitude"
latitude_column_name <- "gps_coordinates _latitude"

# with the actual column names that you've identified for your dataset.
mpox_clean <- mpox_data %>%
  filter(!is.na(.[[longitude_column_name]]) & !is.na(.[[latitude_column_name]]))

# Now, try creating the sf object with the cleaned data
abortion_data_sf <- abortion_data_clean %>%
  st_as_sf(coords = c(longitude_column_name, latitude_column_name), crs = 4326)

# Checking the resulting sf object
print(abortion_data_sf)





# Plotting the map of Isingiro district
ggplot() +
  geom_sf(data = isingiro_district, fill = "lightblue", color = "darkblue") +
  labs(title = "Isingiro District") +
  theme_void() # Use theme_void to remove axes and background for a clean map

# Creating the data frame for the label
centroid_df <- data.frame(x = centroid_coords[,1], y = centroid_coords[,2])

# Adding the label to the plot
ggplot() +
  geom_sf(data = isingiro_district, fill = "lightblue", color = "darkblue") +
  geom_text(data = centroid_df, aes(x = x, y = y, label = "Isingiro District"), size = 5, color = "red") +
  labs(title = "Map of Isingiro District") +
  theme_void()




```

```{r}
## Adding the IgG on the map of Isingiro district

# Installing necessary packages if you haven't already
install.packages("sf")
install.packages("ggplot2")
install.packages("readxl")
install.packages("dplyr")

# Load the packages
library(sf)
library(ggplot2)
library(readxl)
library(dplyr)

# Reading the shapefile for the map from Uganda
shapefile_path <- "~/Documents/R/Reporting of livestock abortions/abortions reporting/uga_admbnda_ubos_20200824_shp/uga_admbnda_adm3_ubos_20200824.shp"
uganda_map <- st_read(shapefile_path)

# Read the dataset with IgG results and GPS coordinates
data_path <- "~/Documents/R/Reporting of livestock abortions/Abortions_all_03102023.xlsx"
abortion_data <- read_excel(data_path)

# Filter out the map for only Isingiro District
isingiro_district <- uganda_map %>% 
  filter(ADM3_EN == "Isingiro") 

# Load necessary libraries
library(readxl)
library(sf)
library(ggplot2)
library(dplyr)

# Setting the path to the Excel file
data_path <- "~/Documents/R/Reporting of livestock abortions/Abortions_all_03102023.xlsx"

# Reading the dataset
abortion_data <- read_excel(data_path)
# Replacing these with the actual column names from your dataframe
longitude_column_name <- "gps_coordinates _longitude"
latitude_column_name <- "gps_coordinates _latitude"

# with the actual column names that you've identified for your dataset.
abortion_data_clean <- abortion_data %>%
  filter(!is.na(.[[longitude_column_name]]) & !is.na(.[[latitude_column_name]]))

# Now, try creating the sf object with the cleaned data
abortion_data_sf <- abortion_data_clean %>%
  st_as_sf(coords = c(longitude_column_name, latitude_column_name), crs = 4326)

# Checking the resulting sf object
print(abortion_data_sf)

```
```{r}
# Replacing these with the actual column names from your dataframe
longitude_column_name <- "gps_coordinates _longitude"
latitude_column_name <- "gps_coordinates _latitude"

# Convert the abortion_data to an sf object using the coordinates
abortion_data_clean <- abortion_data %>%
  filter(!is.na(.[[longitude_column_name]]) & !is.na(.[[latitude_column_name]]))

abortion_data_sf <- abortion_data_clean %>%
  st_as_sf(coords = c(longitude_column_name, latitude_column_name), crs = 4326)

# Ensuring the CRS is the same for both sf objects before plotting
isingiro_district <- st_transform(isingiro_district, crs = st_crs(abortion_data_sf))

# Ploting the map with the rvfelisa_IgG results
ggplot() +
  geom_sf(data = isingiro_district, fill = "lightblue", color = "darkblue") +
  geom_sf(data = abortion_data_sf, aes(color = rvfelisa_IgG), size = 2, alpha = 0.6) +
  scale_color_manual(values = c("Positive" = "red", "Negative" = "green")) +
  labs(title = "IgG Results in Isingiro District", x = "Longitude", y = "Latitude") +
  theme_minimal() +
  theme(legend.position = "right")
```
```{r}
# Calculating the confidence intervals in our prevalence results
install.packages("DescTools")
# Load the DescTools package
library(DescTools)
# Define your outcomes as a matrix or data frame where the first column is the number of successes
# and the second column is the total number of trials
outcomes <- matrix(c(40, 106,
                     13, 66,
                     4, 12,
                     2, 106,
                     1, 12,
                     20, 125,
                     21, 58,
                     9,106,
                     1,66,
                     3,12), ncol = 2, byrow = TRUE)

# Column names for clarity
colnames(outcomes) <- c("Successes", "Total")

# Calculate confidence intervals for each set of outcomes
ci_list <- apply(outcomes, 1, function(x) BinomCI(x[1], x[2]))

# Print the results
print(ci_list)

```

