---
title: "Livestock abortions reporting"
author: "Abel"
date: "03 october 2023"
output: html_document
---

```{r setup, include=FALSE}
#########################################################################################
#Reading the file document
library(readxl)
Abortions_all_03102023 <- read_excel("~/OneDrive - University of Cambridge/R/Reporting of livestock abortions/Abortions reporting/Abortions_all_03102023.xlsx")
View(Abortions_all_03102023)
##############################################################################

## Loading required packages
pacman::p_load(sf,tidyterra,dplyr,tidyverse,arsenal,ggplot2,lubridate,broom,data.table,here,rio,linelist,janitor,knitr,finalfit,forestmodel,kableExtra,skimr,DescTools,lmtest,"epitools")

## Univariate analysis (Supplementary Table 1)
table(Abortions_all_03102023$rvfelisa_IgM)
table(Abortions_all_03102023$rvfelisa_IgG)
table(Abortions_all_03102023$host_rvf)
table(Abortions_all_03102023$host_brucellosis)
table(Abortions_all_03102023$gender)
table(Abortions_all_03102023$occupation)
table(Abortions_all_03102023$education)
table(Abortions_all_03102023$subcounty)
table(Abortions_all_03102023$animalclinical_presentations)
table(Abortions_all_03102023$stage_pregnancy)
table(Abortions_all_03102023$environmental_features)
table(Abortions_all_03102023$history_animal_movement)
table(Abortions_all_03102023$rvf_vaccination_status)
# Calculating the median and IQR for the household
# Remove rows with missing values
cleaned_data <- na.omit(Abortions_all_03102023$household_size)

# Calculate the median of household_size from the cleaned vector
median_value <- median(cleaned_data)

# Calculate the IQR of household_size from the cleaned vector
iqr_value <- IQR(cleaned_data)

# Print the results
cat("Median:", median_value, "\n")
cat("IQR:", iqr_value, "\n")

```

```{r}
#####################################################################################
##Visualing for Table 1 in the manuscript
install.packages("arsenal")
library(arsenal)
Abortion2 <- tableby(rvfelisa_IgM ~ host_rvf + subcounty + rvf_vaccination_status + stage_pregnancy + environmental_features + history_animal_movement, data = Abortions_all_03102023)
summary(Abortion2, text = TRUE)

##Drawing a table to factor in different dates of reporting abortions (17.05.2024) for visualization purposes
library(arsenal)
Abortion2 <- tableby(date_alert_new ~ host_rvf + subcounty + stage_pregnancy + environmental_features + history_animal_movement, data = Abortions_all_03102023) 
summary(Abortion2, text = TRUE)


## Ploting for the results: 25th October 2023
IgG_bivariate <- tableby(rvfelisa_IgG ~ host_rvf, data = Abortions_all_03102023)
summary(IgG_bivariate, text = TRUE)

IgM_bivariate <- tableby(rvfelisa_IgM ~ host_rvf, data = Abortions_all_03102023)
summary(IgM_bivariate, text = TRUE)

#Ploting for brucellosis
brucella <- tableby(brucellosis ~ host_brucellosis, data = Abortions_all_03102023)
summary(brucella, text = TRUE)

#Understanding hosts that suffered both brucellosis and RVF

# Subset the data for hosts with both conditions
hosts_with_both <- Abortions_all_03102023 %>%
  filter(brucellosis == "Positive" & rvfelisa_IgG == "Positive")

# Count how many of each host type have both conditions
count_hosts <- table(hosts_with_both$host_rvf)

# Print the results
print(count_hosts)

# Create a combined variable that indicates the presence of both diseases, one of them, or neither
Abortions_all_03102023$both_diseases <- with(Abortions_all_03102023,
                                             ifelse(brucellosis == "Positive" & rvfelisa_IgG == "Positive", "Both",
                                                    ifelse(brucellosis == "Positive", "Only Brucellosis",
                                                           ifelse(rvfelisa_IgG == "Positive", "Only RVF IgG", "Neither"))))

# Create a contingency table
disease_table <- table(Abortions_all_03102023$host_rvf, Abortions_all_03102023$both_diseases)

# Print the results
print(disease_table)

```
#Univariate analysis before choice of the variables in the final model
```{r}
library(broom)
library(dplyr)

# Define predictor variables for the univariate model for RVF

predictors <- c("host_rvf", "stage_pregnancy", "environmental_features", 
                "history_animal_movement", "number_rvf_vaccinated", 
                "destination_animalmovement", "subcounty")

# Initialize a list to store results
univariate_results <- list()

# Loop through each predictor to fit a univariate model and get standard errors and z-values
for (predictor in predictors) {
    # Fit the univariate model
    formula <- as.formula(paste("rvfelisa_IgG_bin ~", predictor))
    univ_model <- glm(formula, family = binomial, data = Abortions_all_03102023)
    
    univ_summary <- tidy(univ_model, conf.int = TRUE) %>%
        select(term, estimate, std.error, statistic, p.value) %>%  # Select relevant columns
        rename(z_value = statistic) %>%  # Renaming statistic to z_value
        mutate(across(c(std.error, z_value, p.value), round, 3))  # Round to 3 decimal places
    
    univ_summary <- univ_summary %>%
        mutate(Predictor = predictor)
  
    univariate_results[[predictor]] <- univ_summary
}

full_univariate_results <- bind_rows(univariate_results)

full_univariate_results <- full_univariate_results %>%
    select(Predictor, term, estimate, std.error, z_value, p.value)

print(full_univariate_results, n = Inf)  # n = Inf ensures all rows are displayed

options(max.print = 1000)  # Increasing the number of rows that can be printed

# Formating the results to avoid scientific notation
full_univariate_results <- full_univariate_results %>%
    mutate(
        estimate = format(estimate, nsmall = 3, scientific = FALSE),
        std.error = format(std.error, nsmall = 3, scientific = FALSE),
        z_value = format(z_value, nsmall = 3, scientific = FALSE),
        p.value = format(p.value, nsmall = 3, scientific = FALSE)
    )

print(full_univariate_results, row.names = FALSE)

```

# Multivariate analysis for RVF
```{r}
#########################################################################################
library(dplyr)
log_analysis <- Abortions_all_03102023 %>%
  select(date_alert,rvfelisa_IgG) %>% 
  na.omit() %>% 
  mutate(rvfelisa_IgG = if_else(rvfelisa_IgG == "Positive", 1, 0))
logistic.display <- function(model) {
cat("Coefficients:\n")
print(m$coefficients[, 1:4, drop = FALSE])
cat("\n")
cat("Goodness of Fit:\n")
print(c("Deviance = ", round(m$deviance, 3)))
print(c("AIC = ", round(m$aic, 3)))
print(c("BIC = ", round(m$bic, 3)))
}
#count unique values for each variable
sapply(lapply(Abortions_all_03102023, unique), length)
#fiting the positives to be 1 and negatives to be 0
Abortions_all_03102023$rvfelisa_IgG_bin <- ifelse(Abortions_all_03102023$rvfelisa_IgG == "Positive", 1, 0)
# fiting the final logistic regression model

Abortions_all_03102023 <- Abortions_all_03102023 %>% mutate(host_rvf = factor(host_rvf,levels=c('Goats','Cattle','Sheep')))

Abortion4 <- glm(rvfelisa_IgG_bin ~ host_rvf + stage_pregnancy + environmental_features + history_animal_movement, family = binomial, data = Abortions_all_03102023)
summary(Abortion4)

## Separating the date_alert (Target date 14 days)

Abortions_all_03102023$date_alert <- as_datetime(Abortions_all_03102023$date_alert)

cutoff_date <- ymd("2023-02-14") ## 14 days before the set up of call centre (1st March 2023)

before_cutoff <- Abortions_all_03102023 %>% filter(date_alert < cutoff_date) ## Abortions earlier

after_cutoff <- Abortions_all_03102023 %>% filter(date_alert >= cutoff_date) ## Abortions in the window period

number_before <- nrow(before_cutoff)


number_after <- nrow(after_cutoff)

##Creating a new column in the dataset date_alert_new
cutoff_date <- ymd("2023-02-14")
Abortions_all_03102023 <- Abortions_all_03102023 %>%
  mutate(date_alert_new = if_else(date_alert < cutoff_date, "Before 14th Feb 2023", "15th Feb 2023 Onwards"))

## Getting a new dataset without Before 14th Feb 2023 data

updated_Abortions_all_03102023 <- Abortions_all_03102023 %>%
  filter(date_alert_new != "Before 14th Feb 2023")

##Re-running the model just to see in case of any difference

Abortion5 <- glm(rvfelisa_IgG_bin ~ host_rvf + stage_pregnancy + environmental_features + history_animal_movement, family = binomial, data = updated_Abortions_all_03102023)

summary(Abortion5)
```
#Univariate model output for Brucellosis
```{r}
library(broom)
library(dplyr)

# Define predictor variables for the univariate model for RVF

predictors <- c("host_brucellosis", "stage_pregnancy", "environmental_features", 
                "history_animal_movement", "subcounty")

# Initialize a list to store results
brucellosis_univariate_results <- list()

# Loop through each predictor to fit a univariate model and get standard errors and z-values
for (predictor in predictors) {
    # Fit the univariate model
    formula <- as.formula(paste("brucellosis~" , predictor))
    univ_model <- glm(formula, family = binomial, data = Abortions_all_03102023)
    
    univ_summary <- tidy(univ_model, conf.int = TRUE) %>%
        select(term, estimate, std.error, statistic, p.value) %>%  # Select relevant columns
        rename(z_value = statistic) %>%  # Renaming statistic to z_value
        mutate(across(c(std.error, z_value, p.value), round, 3))  # Round to 3 decimal places
    
    univ_summary <- univ_summary %>%
        mutate(Predictor = predictor)
  
    brucellosis_univariate_results[[predictor]] <- univ_summary
}

full_univariate_results <- bind_rows(univariate_results)

full_univariate_results <- full_univariate_results %>%
    select(Predictor, term, estimate, std.error, z_value, p.value)

print(full_univariate_results, n = Inf)  # n = Inf ensures all rows are displayed

```

#Running the logistic regression model for Brucellosis
```{r}
# Updating the brucellosis column to 1 for "Positive" and 0 for "Negative"

Abortions_all_03102023$brucellosis <- ifelse(Abortions_all_03102023$brucellosis == "Positive",1,0)

#Fiting the logistic regression model

Abortions_all_03102023 <- Abortions_all_03102023 %>%
    mutate(host_rvf = relevel(factor(host_brucellosis), ref = "Goats"))

model_brucellosis <- glm(brucellosis ~ host_brucellosis + stage_pregnancy + history_animal_movement, 
                         family = binomial, data = Abortions_all_03102023)

# Check the summary of the model
summary(model_brucellosis)

```

```{r}
#########################################################################################
##Simplifying the tables in R and making the model outputs into a publishable version (Not presented in the final paper)
install.packages("gtsummary")
library(gtsummary)
Abortion4  %>% 
tbl_regression(exponentiate = TRUE) %>% 
bold_p()
########################################################################################
,,,,{r}
```
```{r}
######################################################################################
##Timeliness of Abortions reporting (Fig 2 in the Manuscript)
#Grouping number of abortions per day
IgGresult <- abortions5 %>% select(date_alert,rvfelisa_IgG) %>% na.omit() %>% 
  mutate(rvfelisa_num = if_else(rvfelisa_IgG == "Positive", 1, 0)) %>% 
  group_by(date_alert) %>% 
  summarise(number_reported = n())

# Checking the dates in events are in IgGresult
print(filter(IgGresult, !is.na(event)))
ggplot(IgGresult, aes(x = date_alert, y = number_reported)) +
  geom_line() +
  geom_point() + 
  geom_text(data = subset(IgGresult, !is.na(event)),
            aes(label = event), nudge_y = 5) +  # Nudging the label upwards to avoid overlap
  labs(title = "Number of RVF ELISA IgG Positive Cases Over Time", 
       x = "Date", 
       y = "Number Reported") +
  theme_minimal()
+ ylim(0, max(IgGresult$number_reported) + 10)  # Adjusting the y-axis limits
#####################################################################################
```{r}
#Understanding the duration difference between date_alert and date_sample collection (Appendix: Figure 2)

# Rename the column to remove the extra space
colnames(Abortions_all_03102023)[colnames(Abortions_all_03102023) == "date _samplecollection"] <- "date_samplecollection"

# Convert the corrected column to Date format
Abortions_all_03102023$date_samplecollection <- as.Date(Abortions_all_03102023$date_samplecollection, format = "%Y-%m-%d")
Abortions_all_03102023$date_alert <- as.Date(Abortions_all_03102023$date_alert, format = "%Y-%m-%d")

# Compute the difference in days
Abortions_all_03102023$days_difference <- as.numeric(Abortions_all_03102023$date_samplecollection - Abortions_all_03102023$date_alert)

# Compute the mean and standard deviation of the days difference
average_difference <- mean(Abortions_all_03102023$days_difference, na.rm = TRUE)
sd_difference <- sd(Abortions_all_03102023$days_difference, na.rm = TRUE)

# Print the results
print(paste("Mean:", average_difference))
print(paste("Standard Deviation:", sd_difference))
# Plotting the differences
library(ggplot2)
ggplot(Abortions_all_03102023, aes(x = days_difference)) +
    geom_histogram(binwidth=1, fill="#FF8000", color="black", alpha=2.7) +
    labs(title = "Distribution of days difference between date of alert and date of sample Collection (days)",
         x = "Duration taken to respond to abortion alerts (days)",
         y = "Number of livestock abortions followed up") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5)) +
    coord_cartesian(xlim = c(0, 15))
##########################################################################################
```
```{r}
# Calculating the confidence intervals in our prevalence results (Table I)
# Defining outcomes as a matrix or data frame where the first column is the number of successes and the second column is the total number of trials
outcomes <- matrix(c(40, 106,
                     13, 66,
                     4, 12,
                     2, 106,
                     1, 12,
                     20, 125,
                     21, 58,
                     9,106,
                     1,66,
                     3,12,
                     0,66,
                     0,16), ncol = 2, byrow = TRUE)

# Column names for clarity
colnames(outcomes) <- c("Successes", "Total")

# Calculate confidence intervals for each set of outcomes
ci_list <- apply(outcomes, 1, function(x) BinomCI(x[1], x[2]))
print(ci_list)
##############################################################################################
```

