---
title: "Livestock abortions reporting"
author: "Abel"
date: "03 october 2023"
output: html_document
---

```{r setup, include=FALSE}
#########################################################################################
#Reading the file document
library(readxl)
Abortions_all_03102023 <- read_excel("~/OneDrive - University of Cambridge/R/Reporting of livestock abortions/Abortions reporting/Abortions_all_03102023.xlsx")
View(Abortions_all_03102023)
########################################################################################
```
```{r}

```
```{r}
## Loading required packages
pacman::p_load(sf,tidyterra,dplyr,tidyverse,arsenal,ggplot2,lubridate,broom,data.table,here,rio,linelist,janitor,knitr,finalfit,forestmodel,kableExtra,skimr,DescTools,lmtest,"epitools")

## Univariate analysis (Supplementary Table 1)
table(Abortions_all_03102023$rvfelisa_IgM)
table(Abortions_all_03102023$rvfelisa_IgG)
table(Abortions_all_03102023$host_rvf)
table(Abortions_all_03102023$host_brucellosis)
table(Abortions_all_03102023$gender)
table(Abortions_all_03102023$occupation)
table(Abortions_all_03102023$education)
table(Abortions_all_03102023$subcounty)
table(Abortions_all_03102023$animalclinical_presentations)
table(Abortions_all_03102023$stage_pregnancy)
table(Abortions_all_03102023$environmental_features)
table(Abortions_all_03102023$history_animal_movement)
table(Abortions_all_03102023$rvf_vaccination_status)
# Calculating the median and IQR for the household
# Remove rows with missing values
cleaned_data <- na.omit(Abortions_all_03102023$household_size)

# Calculate the median of household_size from the cleaned vector
median_value <- median(cleaned_data)

# Calculate the IQR of household_size from the cleaned vector
iqr_value <- IQR(cleaned_data)

# Print the results
cat("Median:", median_value, "\n")
cat("IQR:", iqr_value, "\n")

```

```{r}
#####################################################################################
##Visualing for Table 1 in the manuscript
install.packages("arsenal")
library(arsenal)
Abortion2 <- tableby(rvfelisa_IgM ~ host_rvf + subcounty + rvf_vaccination_status + stage_pregnancy + environmental_features + history_animal_movement, data = Abortions_all_03102023)
summary(Abortion2, text = TRUE)

##Drawing a table to factor in different dates of reporting abortions (17.05.2024) for visualization purposes
library(arsenal)
Abortion2 <- tableby(date_alert_new ~ host_rvf + subcounty + stage_pregnancy + environmental_features + history_animal_movement, data = Abortions_all_03102023) 
summary(Abortion2, text = TRUE)


## Ploting for the results: 25th October 2023
IgG_bivariate <- tableby(rvfelisa_IgG ~ host_rvf, data = Abortions_all_03102023)
summary(IgG_bivariate, text = TRUE)

IgM_bivariate <- tableby(rvfelisa_IgM ~ host_rvf, data = Abortions_all_03102023)
summary(IgM_bivariate, text = TRUE)

#Ploting for brucellosis
brucella<- tableby(brucellosis ~ host_brucellosis, data = Abortions_all_03102023)
summary(brucella, text = TRUE)

#Understanding hosts that suffered both brucellosis and RVF

# Subset the data for hosts with both conditions
hosts_with_both <- Abortions_all_03102023 %>%
  filter(brucellosis == "Positive" & rvfelisa_IgG == "Positive")

# Count how many of each host type have both conditions
count_hosts <- table(hosts_with_both$host_rvf)

# Print the results
print(count_hosts)

# Create a combined variable that indicates the presence of both diseases, one of them, or neither
Abortions_all_03102023$both_diseases <- with(Abortions_all_03102023,
                                             ifelse(brucellosis == "Positive" & rvfelisa_IgG == "Positive", "Both",
                                                    ifelse(brucellosis == "Positive", "Only Brucellosis",
                                                           ifelse(rvfelisa_IgG == "Positive", "Only RVF IgG", "Neither"))))

# Create a contingency table
disease_table <- table(Abortions_all_03102023$host_rvf, Abortions_all_03102023$both_diseases)

# Print the results
print(disease_table)

```
```{r}
#########################################################################################
# Multivariate analysis
library(dplyr)

log_analysis <- Abortions_all_03102023 %>%
  select(date_alert,rvfelisa_IgG) %>% 
  na.omit() %>% 
  mutate(rvfelisa_IgG = if_else(rvfelisa_IgG == "Positive", 1, 0))
logistic.display <- function(model) {
cat("Coefficients:\n")
print(m$coefficients[, 1:4, drop = FALSE])
cat("\n")
cat("Goodness of Fit:\n")
print(c("Deviance = ", round(m$deviance, 3)))
print(c("AIC = ", round(m$aic, 3)))
print(c("BIC = ", round(m$bic, 3)))
}
#count unique values for each variable
sapply(lapply(Abortions_all_03102023, unique), length)
#fiting the positives to be 1 and negatives to be 0
Abortions_all_03102023$rvfelisa_IgG_bin <- ifelse(Abortions_all_03102023$rvfelisa_IgG == "Positive", 1, 0)
# fiting the final logistic regression model

Abortions_all_03102023 <- Abortions_all_03102023 %>% mutate(host_rvf = factor(host_rvf,levels=c('Goats','Cattle','Sheep')))

Abortion4 <- glm(rvfelisa_IgG_bin ~ host_rvf + stage_pregnancy + environmental_features + history_animal_movement, family = binomial, data = Abortions_all_03102023)
summary(Abortion4)

## Separating the date_alert (Target date 14 days)

Abortions_all_03102023$date_alert <- as_datetime(Abortions_all_03102023$date_alert)

cutoff_date <- ymd("2023-02-14") ## 14 days before the set up of call centre (1st March 2023)

before_cutoff <- Abortions_all_03102023 %>% filter(date_alert < cutoff_date) ## Abortions earlier

after_cutoff <- Abortions_all_03102023 %>% filter(date_alert >= cutoff_date) ## Abortions in the window period

number_before <- nrow(before_cutoff)


number_after <- nrow(after_cutoff)

##Creating a new column in the dataset date_alert_new
cutoff_date <- ymd("2023-02-14")
Abortions_all_03102023 <- Abortions_all_03102023 %>%
  mutate(date_alert_new = if_else(date_alert < cutoff_date, "Before 14th Feb 2023", "15th Feb 2023 Onwards"))

## Getting a new dataset without Before 14th Feb 2023 data

updated_Abortions_all_03102023 <- Abortions_all_03102023 %>%
  filter(date_alert_new != "Before 14th Feb 2023")

##Re-running the model just to see in case of any difference

Abortion5 <- glm(rvfelisa_IgG_bin ~ host_rvf + stage_pregnancy + environmental_features + history_animal_movement, family = binomial, data = updated_Abortions_all_03102023)

summary(Abortion5)
#########################################################################################
##Simplifying the tables in R and making the model outputs into a publishable version (Not presented in the final paper)
install.packages("gtsummary")
library(gtsummary)
Abortion4  %>% 
tbl_regression(exponentiate = TRUE) %>% 
bold_p()
########################################################################################
,,,,{r}
```
```{r}
######################################################################################
##Timeliness of Abortions reporting (Fig 2 in the Manuscript)
#Grouping number of abortions per day
IgGresult <- abortions5 %>% select(date_alert,rvfelisa_IgG) %>% na.omit() %>% 
  mutate(rvfelisa_num = if_else(rvfelisa_IgG == "Positive", 1, 0)) %>% 
  group_by(date_alert) %>% 
  summarise(number_reported = n())

# Checking the dates in events are in IgGresult
print(filter(IgGresult, !is.na(event)))
ggplot(IgGresult, aes(x = date_alert, y = number_reported)) +
  geom_line() +
  geom_point() + 
  geom_text(data = subset(IgGresult, !is.na(event)),
            aes(label = event), nudge_y = 5) +  # Nudging the label upwards to avoid overlap
  labs(title = "Number of RVF ELISA IgG Positive Cases Over Time", 
       x = "Date", 
       y = "Number Reported") +
  theme_minimal()
+ ylim(0, max(IgGresult$number_reported) + 10)  # Adjusting the y-axis limits
#####################################################################################
```{r}
#Understanding the duration difference between date_alert and date_sample collection (Appendix: Figure 2)

Abortions_all_03102023[["date_alert"]] <- as.Date(Abortions_all_03102023[["date_alert"]], format = "%Y-%m-%d") 
Abortions_all_03102023[["date _samplecollection"]] <- as.Date(Abortions_all_03102023[["date _samplecollection"]], format = "%Y-%m-%d")
Abortions_all_03102023$days_difference <- as.numeric(Abortions_all_03102023[["date _samplecollection"]] - Abortions_all_03102023[["date_alert"]])
average_difference <- mean(Abortions_all_03102023$days_difference, na.rm = TRUE) # Getting the difference between the dates
print(average_difference)

# Plotting the differences
library(ggplot2)
ggplot(Abortions_all_03102023, aes(x = days_difference)) +
    geom_histogram(binwidth=1, fill="#FF8000", color="black", alpha=0.7) +
    labs(title = "Distribution of Days Difference between Alert and Sample Collection",
         x = "Duration to respond to abortion alerts",
         y = "Number of abortions followed up") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5)) +
    coord_cartesian(xlim = c(0, 15))  
##########################################################################################
```
```{r}
# Calculating the confidence intervals in our prevalence results (Table I)
# Defining outcomes as a matrix or data frame where the first column is the number of successes and the second column is the total number of trials
outcomes <- matrix(c(40, 106,
                     13, 66,
                     4, 12,
                     2, 106,
                     1, 12,
                     20, 125,
                     21, 58,
                     9,106,
                     1,66,
                     3,12), ncol = 2, byrow = TRUE)

# Column names for clarity
colnames(outcomes) <- c("Successes", "Total")

# Calculate confidence intervals for each set of outcomes
ci_list <- apply(outcomes, 1, function(x) BinomCI(x[1], x[2]))
print(ci_list)
##############################################################################################
```

